---
title: "survival-associated immune-related hub genes"
date: "Created: 2021-09-01 Updated: `r Sys.Date()`"
output: 
  html_notebook:
    codes: hide
---

### Introdution
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(convert)
library(ggplot2)
library(survival)
library(survminer)
library(data.table)
library(maxstat)

# rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)

grp <- c("Normal", "Tumor")
grp.col <- c("#568875", "#73FAFC")
```


### Importing data 
```{r}
HNSC_ExprSet_Immune <- readRDS("../../Result/Profile/HNSC_ExprSet_ImmuneGene.RDS")
HNSC_ExprSet_Immune

HubGene <- fread("../../Result/WGCNA/CytoscapeInput-nodes-brown-yellow-degrees.csv")
```


### survival-associated immune-related hub genes
```{r}
get_SurHubGene <- function(dataset=HNSC_ExprSet_Immune,
                           genelist=HubGene,
                           ntop=50,
                           Pval=0.05){

  # dataset=HNSC_ExprSet_Immune  
  # genelist=HubGene
  # ntop=50
  # Pval=0.2
  
  gene_top <- genelist %>% arrange(desc(degree.layout)) %>%
    dplyr::slice(1:ntop) %>%
    dplyr::select(name, degree.layout) %>%
    setNames(c("GeneID", "Degree"))
  
  profile <- exprs(dataset) %>% data.frame()
  phenotype <- pData(dataset)
  
  prof <- profile[gene_top$GeneID, ]
  
  mdat <- phenotype %>% rownames_to_column("Barcode") %>%
    inner_join(t(prof) %>% data.frame() %>% rownames_to_column("Barcode"),
               by = "Barcode") %>%
    dplyr::select(-SampleID) %>%
    filter(Group!=grp[1]) %>% 
    column_to_rownames("Barcode") 
  
  datphe <- mdat %>% dplyr::select(OS.Time, OS) 
  datprf <- mdat %>% dplyr::select(rownames(prof)) 
  
  sur_res <- apply(datprf, 2, function(x, y){ 
    dat <- cbind(y, Value=x)
    #dat <- cbind(datphe, Value=datprf$ESRRG)
    fit <- maxstat.test(Surv(OS.Time, OS) ~ Value, 
                        data = dat, 
                        smethod = "LogRank", 
                        pmethod = "Lau92") 
    statistic <- as.numeric(fit$statistic)
    pvalue <- as.numeric(fit$p.value)
    cutpoint <- as.numeric(fit$estimate)
    res <- c(statistic, pvalue, cutpoint)
    return(res)
    
    }, datphe) %>% t() %>% data.frame() %>%
    setNames(c("Statistic", "P.value", "Estimated_cutpoint")) %>%
    rownames_to_column("GeneID") 
  
  res <- sur_res %>% inner_join(gene_top, by = "GeneID") %>%
    mutate(Status=ifelse(P.value < 0.2, "Remain", "Discard"))
  
  return(res)
}


if(!dir.exists("../../Result/Survival")){
  dir.create("../../Result/Survival", recursive = T)
}

survival_hubgene <- get_SurHubGene(dataset=HNSC_ExprSet_Immune, Pval=0.2)
write.csv(survival_hubgene, "../../Result/Survival/HNSC_SurHubGene.csv", row.names = F)
DT::datatable(survival_hubgene)
```


### Optimal cut-off value for each hub gene
```{r}
get_cutpoint <- function(dataset=HNSC_ExprSet_Immune,
                         genelist=HubGene,
                         ntop=50){

  # dataset=HNSC_ExprSet_Immune
  # genelist=HubGene
  # ntop=50
  
  gene_top <- genelist %>% arrange(desc(degree.layout)) %>%
    dplyr::slice(1:ntop) %>%
    dplyr::select(name, degree.layout) %>%
    setNames(c("GeneID", "Degree"))
  
  profile <- exprs(dataset) %>% data.frame()
  phenotype <- pData(dataset)
  
  prof <- profile[gene_top$GeneID, ]
  
  mdat <- phenotype %>% rownames_to_column("Barcode") %>%
    inner_join(t(prof) %>% data.frame() %>% rownames_to_column("Barcode"),
               by = "Barcode") %>%
    dplyr::select(-SampleID) %>%
    filter(Group!=grp[1]) %>% 
    column_to_rownames("Barcode") 
  
  datphe <- mdat %>% dplyr::select(OS.Time, OS)
  datprf <- mdat %>% dplyr::select(rownames(prof))
  
  covariables <- colnames(datprf)
  res_threshold <- lapply(covariables, function(x){
    fit <- surv_cutpoint(data = mdat,
                         time = "OS.Time",
                         event = "OS",
                         variables = x)
    res <- as.numeric(summary(fit))
    return(res)
  }) %>% data.frame() %>% t() %>%
    data.frame() %>%
    setNames(c("cutpoint", "statistic"))
  rownames(res_threshold) <- covariables
  
  
  res <- res_threshold %>% rownames_to_column("GeneID")
  
  return(res)
}

if(!dir.exists("../../Result/Survival")){
  dir.create("../../Result/Survival", recursive = T)
}

hubgene_cutoff <- get_cutpoint(dataset=HNSC_ExprSet_Immune)
DT::datatable(hubgene_cutoff)
```


**Notes: the values of estimated_cutpoint in survival_hubgene equal to the values of cutpoint in hubgene_cutoff**


### systemic information
```{r}
sessionInfo()
```


### Reference

1. [Maximally Selected Rank and Statistics](https://www.rdocumentation.org/packages/maxstat/versions/0.7-24/topics/maxstat.test)

2. [Determine optimal cutpoints for numerical variables in survival plots](http://r-addict.com/2016/11/21/Optimal-Cutpoint-maxstat.html)
