---
title: "IRGPI: Construction and validation of the immune-related gene prognostic index"
date: "Created: 2021-09-02 Updated: `r Sys.Date()`"
output: 
  html_notebook:
    codes: hide
---

### Introdution

The IRGPI of each sample was calculated by multiplying the expression values of certain genes by their weight in the Cox model and then adding them together.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(convert)
library(data.table)
library(ggplot2)
library(survival)

# rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)

grp <- c("Normal", "Tumor")
grp.col <- c("#568875", "#73FAFC")
```


### Importing Data 
```{r}
HNSC_ExprSet_Immune <- readRDS("../../Result/Profile/HNSC_ExprSet_ImmuneGene.RDS")
HNSC_ExprSet_Immune

HNSC_SurHubGene <- fread("../../Result/Survival/HNSC_SurHubGene.csv")

HNSC_clinical <- fread("../../Study/SeqData/TCGA_HNSC/Clinical/TCGA-HNSC-post_clinical.csv")
```


### Cox Proportional-Hazards Model & Prognosic index
```{r}
Cox_ProIndex <- function(dataset=HNSC_ExprSet_Immune,
                    datagene=HNSC_SurHubGene,
                    metadata=HNSC_clinical,
                    Pval=0.1){

  # dataset=HNSC_ExprSet_Immune
  # datagene=HNSC_SurHubGene
  # Pval=0.1
  # metadata=HNSC_clinical

  phen <- pData(dataset)
  prof <- exprs(dataset)
  feature <- datagene %>% filter(Status != "Discard")

  fid <- intersect(rownames(prof), feature$GeneID)
  prof_cln <- prof[rownames(prof)%in%fid, ] %>%
    t() %>% data.frame() %>%
    rownames_to_column("Barcode")

  mdat <- phen %>% rownames_to_column("Barcode") %>%
    filter(Group != grp[1]) %>%
    dplyr::select(Barcode, OS, OS.Time) %>%
    inner_join(prof_cln, by = "Barcode") %>%
    mutate(Barcode=gsub("_01A|_01B|_06A", "", Barcode))
  
  # deduplicated samples
  pid <- intersect(metadata$Barcode, mdat$Barcode)
  mdat_cln <- mdat[pmatch(pid, mdat$Barcode), ]
  clinical_cln <- metadata[pmatch(pid, metadata$Barcode), ]
  print(dim(mdat_cln))
  
  datphe <- mdat_cln %>% dplyr::select(OS.Time, OS) 
  datprf <- mdat_cln %>% dplyr::select(colnames(prof_cln))
  rownames(datprf) <- NULL

  covariates <- colnames(datprf)[-1]
  univ_formulas <- sapply(covariates, function(x){as.formula(paste("Surv(OS.Time, OS)~", x))})
  univ_models <- lapply(univ_formulas, function(x){coxph(x, data = mdat_cln)})

  # Extract data
  univ_results <- lapply(univ_models,
                         function(x){
                            x <- summary(x)
                            p.value <- signif(x$wald["pvalue"], digits=4)
                            wald.test <- signif(x$wald["test"], digits=4)
                            beta <- signif(x$coef[1], digits=4) #coeficient beta
                            HR <-signif(x$coef[2], digits=2) #exp(beta)
                            HR_lower <- signif(x$conf.int[, "lower .95"], 2)
                            HR_upper <- signif(x$conf.int[, "upper .95"], 2)
                            HR_95 <- paste0(HR, " (", HR_lower, "-", HR_upper, ")")
                            res <- c(beta, HR_95, HR, HR_lower, HR_upper, wald.test, p.value)
                            names(res) <- c("beta", "HR(95% CI)", "HR_mean", 
                                            "HR_lower", "HR_upper",
                                            "wald.test", "P.Value")
                            return(res)
                           })
  res_cox <- t(as.data.frame(univ_results, check.names = FALSE)) %>%
    data.frame() %>%
    rownames_to_column("GeneID") 
  
  # prognostic index
  gene_cox <- res_cox %>% 
    mutate(P.Value=as.numeric(P.Value),
           beta=as.numeric(beta)) %>%
    filter(as.numeric(P.Value) < Pval) %>%
    dplyr::select(GeneID, beta) %>%
    column_to_rownames("GeneID") %>%
    as.matrix()
  
  prf_index <- datprf %>% column_to_rownames("Barcode") %>%
    dplyr::select(rownames(gene_cox)) %>%
    as.matrix()
  
  Pindex <- prf_index %*% gene_cox %>%
    data.frame() %>% setNames("Prognostic_Score") %>%
    rownames_to_column("Barcode") %>%
    inner_join(clinical_cln, by="Barcode")
  
  # Optimal cut-off value for prognostic index
  Pindex_fit <- surv_cutpoint(data = Pindex,
                         time = "OS.Time",
                         event = "OS",
                         variables = "Prognostic_Score")
  Pindex_threshold <- as.numeric(summary(Pindex_fit))[1]
  Pindex$IRGPI_group <- ifelse(Pindex$Prognostic_Score > Pindex_threshold, "IRGPI_high", "IRGPI_low")
  
  
  # survival-associated immune-related hub genes with low and high cluster
  sur_gene <- clinical_cln %>% 
    inner_join(data.frame(prf_index) %>% 
                 rownames_to_column("Barcode"), 
               by = "Barcode") %>%
    data.frame()
  sur_gene_covariables <- rownames(gene_cox)
  sur_gene__threshold <- lapply(sur_gene_covariables, function(x){
    fit <- surv_cutpoint(data = sur_gene,
                         time = "OS.Time",
                         event = "OS",
                         variables = x)
    res <- as.numeric(summary(fit))
    return(res)
  }) %>% data.frame() %>% t() %>%
    data.frame() %>%
    setNames(c("cutpoint", "statistic"))
  rownames(sur_gene__threshold) <- sur_gene_covariables  
  df_group <- data.frame(matrix(NA, ncol=length(sur_gene_covariables), nrow=nrow(sur_gene)))
  rownames(df_group) <- sur_gene$Barcode
  colnames(df_group) <- paste0(sur_gene_covariables, "_group")
  for(i in 1:length(sur_gene_covariables)){
    gene_index <- which(colnames(sur_gene) == sur_gene_covariables[i])
    gene_group <- paste0(sur_gene_covariables[i], c("_high", "_low"))
    df_group[, i] <- ifelse(sur_gene[, gene_index] > sur_gene__threshold$cutpoint[i], 
                            gene_group[2], gene_group[1])
  }
  
  sur_gene_res <- inner_join(sur_gene, df_group %>% rownames_to_column("Barcode"),
                             by = "Barcode")
  
  res <- list(cox=res_cox,
              pindex=Pindex,
              surprf=sur_gene_res)

  return(res)
}

if(!dir.exists("../../Result/Survival")){
  dir.create("../../Result/Survival", recursive = T)
}

Prognostic_Index <- Cox_ProIndex(
                    dataset=HNSC_ExprSet_Immune,
                    datagene=HNSC_SurHubGene,
                    metadata=HNSC_clinical,
                    Pval=0.1)

write.csv(Prognostic_Index$cox, "../../Result/Survival/HNSC_SurHubGene_cox.csv", row.names = F)
write.csv(Prognostic_Index$pindex, "../../Result/Survival/HNSC_SurHubGene_Pindex.csv", row.names = F)
write.csv(Prognostic_Index$surprf, "../../Result/Survival/HNSC_SurHubGene_profile.csv", row.names = F)
DT::datatable(Prognostic_Index$cox)
DT::datatable(Prognostic_Index$pindex)
```


### Univariate Cox analysis of 21 immune-related hub genes
```{r, fig.width=6, fig.height=8}
library(forestplot)
tabletext <- cbind(c(NA, "Hub genes", Prognostic_Index$cox$GeneID), 
                   c(NA, "P-value", as.numeric(Prognostic_Index$cox$P.Value)),
                   c(NA, "HR(95% CI)", Prognostic_Index$cox$HR.95..CI.))
forestplot(labeltext=tabletext, 
           graph.pos=2,                     
           col=fpColors(box="#D55E00", lines="#CC79A7", zero="gray50"),
           mean=c(NA, NA, as.numeric(Prognostic_Index$cox$HR_mean)),
           lower=c(NA, NA, as.numeric(Prognostic_Index$cox$HR_lower)),
           upper=c(NA, NA, as.numeric(Prognostic_Index$cox$HR_upper)),
           boxsize=0.3,
           lwd.ci=2,            
           ci.vertices.height=0.08,
           ci.vertices=TRUE,
           zero=1,
           lwd.zero=1,      
           colgap=unit(5, "mm"),  
           xticks=c(0.5, 1, 1.5), 
           lwd.xaxis=1,          
           lineheight=unit(.8, "cm"), 
           graphwidth=unit(.3, "npc"), 
           cex=0.9, 
           fn.ci_norm=fpDrawCircleCI,           
           hrzl_lines=list("3" =gpar(lwd=2, col="black"),  
                           "24"=gpar(lwd=2, col="black")),
           mar=unit(rep(0.5, times=4), "cm"),             
           txt_gp=fpTxtGp(label=gpar(cex=1),                
                          ticks=gpar(cex=1.5),
                          xlab=gpar(cex=1.25),
                          title=gpar(cex=1.2)),
           xlab="Hazard Ratio")
```

**Notes: the three genes were SPINK5 NR3C2 AZGP1**


### Kaplan-Meier survival analysis of the three genes significant in the univariate Cox analysis
```{r}
get_surplot <- function(dataset=Prognostic_Index$surprf, 
                        group_name="SPINK5"){
  
  # dataset=Prognostic_Index$surprf
  # group_name="SPINK5"
  
  # pvalue
  colnames(dataset)[which(colnames(dataset) == paste0(group_name, "_group"))] <- "group_info"
  print(table(dataset$group_info))
  
  dat_sur <- dataset %>% dplyr::select(OS.Time, OS, group_info) %>%
    mutate(group_info=factor(group_info, levels = paste0(group_name, c("_high", "_low")))) 
    
  factors <- levels(dat_sur$group_info)
  
  cox <- coxph(Surv(OS.Time, OS) ~ group_info, data = dat_sur)
  tmp <- summary(cox)
  tmp.wald <- data.frame(t(tmp$waldtest)) %>%
        setNames(c("Wald_test", "Wald_df", "Wald_pvlaue"))
  tmp.lg <- data.frame(t(tmp$logtest)) %>%
        setNames(c("lg_rank", "lg_rank_df", "lg_rank_pvlaue"))
  tmp.total <- cbind(tmp.wald, tmp.lg) 
  
  pvalue <- paste(paste0("Log-Rank P=", signif(tmp.lg$lg_rank_pvlaue, 3)), 
                  paste0("Cox P=", signif(tmp.wald$Wald_pvlaue, 3)), sep = "\n")
  
  # plot
  fit <- survfit(Surv(OS.Time, OS) ~ group_info, data = dat_sur)
  info <- data.frame(time = fit$time,
                  n.risk = fit$n.risk,
                  n.event = fit$n.event,
                  n.censor = fit$n.censor,
                  surv = fit$surv,
                  upper = fit$upper,
                  lower = fit$lower)
  
  pl <- ggsurvplot(fit,
             data = dat_sur,  
             surv.median.line = "hv",
             add.all = FALSE,
             palette = c("red", "blue"), #"aaas",
             risk.table = TRUE, 
             xlab = "Follow up time(Years)", 
             legend = c(0.9, 0.4), 
             legend.title = "",
             legend.labs = factors, #c("all", factors), 
             break.x.by = 2,
             font.legend = c(10, "italic"),
             ggtheme = theme_bw())
  pl$plot <- pl$plot + 
             annotate("text", x=3, y=0.2, label=pvalue)
  
  res <- list(info=info, pl=pl)
  return(res)
}
```



#### SPINK5
```{r, fig.width=8, fig.height=6}
get_surplot(dataset=Prognostic_Index$surprf, 
            group_name="SPINK5")
```


#### NR3C2 
```{r, fig.width=8, fig.height=6}
get_surplot(dataset=Prognostic_Index$surprf, 
            group_name="NR3C2")
```


#### AZGP1
```{r, fig.width=8, fig.height=6}
get_surplot(dataset=Prognostic_Index$surprf, 
            group_name="AZGP1")
```


#### Prognostic score
```{r, fig.width=8, fig.height=6}
get_surplot(dataset=Prognostic_Index$pindex, 
            group_name="IRGPI")
```


### systemic information
```{r}
sessionInfo()
```


### Reference

1. [使用forestplot绘制森林图](https://cloud.tencent.com/developer/article/1654977)
