---
title: "Somatic mutations of immune-related hub genes"
date: "Created: 2021-09-14 Updated: `r Sys.Date()`"
output: 
  html_notebook:
    codes: hide
---

### Introdution

To reveal relevant genetic alterations, somatic mutations of 22 immune-related hub genes were analyzed by using the ComplexHeatmap package of R.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(convert)
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(maftools)

# rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)

grp <- c("Normal", "Tumor")
grp.col <- c("#568875", "#73FAFC")
```


### Importing Data 
```{r}
phenotype <- read.csv("../../Study/SeqData/TCGA_HNSC/Clinical/TCGA-HNSC-post_clinical.csv")
SurHubGene <- fread("../../Result/Survival/HNSC_SurHubGene.csv")
SomaticMutation <- fread("../../Study/SeqData/TCGA_HNSC/Somatic_Mutation/TCGA-HNSC_mutect_somatic_maf.tsv")
```


### Filtering Gene by Survival hub genes
```{r}
dat <- SomaticMutation %>% filter(Hugo_Symbol%in%SurHubGene$GeneID) 
datlam <- read.maf(maf = SomaticMutation)
```


### whole picture of somatic mutation
```{r}
plotmafSummary(maf = datlam, rmOutlier = TRUE, addStat = 'median', dashboard = TRUE, titvRaw = FALSE)
```


### oncoplot
```{r, fig.width=10, fig.height=8}
oncoplot(maf = datlam, top = 20)
```


### interaction among genes
```{r}
Interact <- somaticInteractions(maf = datlam, top = 20, pvalue = c(0.05, 0.1))
Interact$gene_sets
```


### systemic information
```{r}
sessionInfo()
```


### Reference

1. [maftools|TCGA肿瘤突变数据的汇总，分析和可视化](https://mp.weixin.qq.com/s?__biz=MzIyNDI1MzgzOQ==&mid=2650394616&idx=2&sn=92f512cef2a262a12eaec7992107ec03&chksm=f01cad18c76b240e69d50a64d20e20a65ed5926fcd838d2ad0dffb9549a62b172951a9cf4eec&scene=21#wechat_redirect)

2. [maftools | 从头开始绘制发表级oncoplot（瀑布图）](https://mp.weixin.qq.com/s?__biz=MzIyNDI1MzgzOQ==&mid=2650394240&idx=1&sn=44b7769942c63fb94c2547a5ad5bb771&chksm=f01cac60c76b25766587c8af2338d06dc39646479d748f5a33ea2bb588251f719b21118a6c17&scene=21#wechat_redirect)
